---
title: "Data Heatmaps"
author: "Nick R. Bachelder"
date: "2/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(igraph)
library(ggplot2)
library(tidyverse)
library(viridis)
library(plyr)

options(max.print=1000000)
interactions_b <- read.csv('interactions_dup.csv') %>% select(-X)

interactions_b[c(1,2,3,4,5),] 
```


DO FOR SPECIFICALLY "adrenidae"
```{r}
for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}

for (c in colnames(interactions_b)) {
  interactions_b[[c]] <-  tolower(interactions_b[[c]])
}

interactions_andrenidae <- interactions_b %>% filter(sourceTaxonFamilyName == 'andrenidae')
```


Heatmap for Recorded Ineractions of Family of Bee and Plantae Families.
```{r}
beeCite2 <- interactions_andrenidae %>% dplyr::select(c(sourceTaxonFamilyName, targetTaxonFamilyName)) %>%
  dplyr::rename(interestBee = sourceTaxonFamilyName, interact = targetTaxonFamilyName) %>%
  group_by(interestBee, interact) %>% count() %>% filter(freq >= 300) %>% ungroup() %>% complete(interestBee, interact, fill = list(freq = 0))


beeCite2 %>% ggplot( aes(x=interact, y= freq)) +
    geom_segment( aes(x=interact ,xend=interact, y=0, yend= freq), color="black") +
    geom_point(size=4, color="#69b3a2") +
    coord_flip() + xlab("Plantae Genus") + ylab('Number of Interactions') + ggtitle('Ineractions of Andrenidae Bee Family and Various Plantae Families')

beeCite2
  
```


Heatmap for Recorded Ineractions of Genus of Bee and Plantae Genus
```{r}
beeCite2 <- interactions_andrenidae %>% dplyr::select(c(sourceTaxonGenusName, targetTaxonGenusName)) %>% dplyr::rename(interestBee = sourceTaxonGenusName, interact = targetTaxonGenusName) %>%
  group_by(interestBee, interact) %>% count() %>% filter(freq >=400) %>% ungroup() %>% complete(interestBee, interact, fill = list(freq = 0))


ggplot(beeCite2, aes(interestBee, interact, fill= freq)) +
  geom_tile(color = 'black') + xlab(paste('Genus of Bee')) + labs(fill = "Number of Citations") + ylab(paste('Plant Genus Bees Interact With')) + 
  ggtitle(paste('Recorded Ineractions of Genus of Andrenidae Bee and Plantae Genus')) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Heatmap for Recorded Ineractions of Family of Bee and Plantae Genus
```{r}
beeCite2 <- interactions_andrenidae %>% dplyr::select(c(sourceTaxonFamilyName, targetTaxonGenusName)) %>% dplyr::rename(interestBee = sourceTaxonFamilyName, interact = targetTaxonGenusName) %>%
  group_by(interestBee, interact) %>% count() %>% filter(freq>=500) %>% ungroup() %>% complete(interestBee, interact, fill = list(freq = 0))

beeCite2 %>% ggplot( aes(x=interact, y=freq)) +
    geom_segment( aes(x=interact ,xend=interact, y=0, yend= freq), color="black") +
    geom_point(size=4, color="#69b3a2") +
    coord_flip() + xlab("Plantae Genus") + ylab('Number of Interactions') + ggtitle('Ineractions of Andrenidae Bee Family and Plantae Genus')
  
```

Heatmap for Recorded Ineractions of Genus of Bee and Plantae Family
```{r}
beeCite2 <- interactions_andrenidae %>% dplyr::select(c(sourceTaxonGenusName, targetTaxonFamilyName)) %>% dplyr::rename(interestBee = sourceTaxonGenusName, interact = targetTaxonFamilyName) %>%
  group_by(interestBee, interact) %>% count() %>% filter(freq>=600) %>% ungroup() %>% complete(interestBee, interact, fill = list(freq = 0))


ggplot(beeCite2, aes(interestBee, interact, fill= freq)) +
  geom_tile(color = 'black') + xlab(paste('Genus of Bee')) + labs(fill = "Number of Citations") + ylab(paste('Plant Family Bees Interact With')) + 
  ggtitle(paste('Recorded Ineractions of Genus of Andrenidae Bee and Plantae Families')) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



#### WEEK 9 : FIX Heatmaps

I determined that the heatmaps are slightly flawed in the sense that, when the filter is applied, it wipes all the relationships below the filter. Then, subsequently, it refills these relationships with 0 if the bee level has another plant relationship that is above the threshold. Therefore, there are values of bee relationships that are listed as 0, but should be somewhere between 0 and the filter level. Here is my solution:


Here is a flawed heatmap for family to family. Many of these 0s are slightly higher than 0.



```{r}
ggplot(beeCite2, aes(interestBee, interact, fill= freq)) +
  geom_tile(color = 'black') + xlab(paste('Genus of Bee')) + labs(fill = "Number of Citations") + ylab(paste('Plant Family Bees Interact With')) + 
  ggtitle(paste('Recorded Ineractions of Family of Bee and Plantae Families')) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+ scale_fill_gradient(
    low = "white",
    high = "blue",
    guide = "colourbar"
  ) + geom_text(aes(label = freq))
```

